### 2018-01-24 parallel-tractability-J-v9/response.tex
added review response file to keep track of fixed issues and the according reply

\documentclass{article}
\usepackage{amsmath}
\usepackage{url}

\let\quoteOld\quote
\let\endquoteOld\endquote
\renewenvironment{quote}{\quoteOld\itshape}{\endquoteOld}

% use the following command line to convert to standard text file
% pandoc reviews.tex -o rebuttal.txt
\setlength{\parindent}{0pt}

\begin{document}

We thank the reviewers for the time and effort invested in the detailed reviews and the valuable suggestions and comments to improve the quality of the paper.  Below we reply to the individual points of the reviewers.

\section{Rebuttal to Review 1}

\begin{quote}
There should, however, be some discussion somewhere in the paper on the
relationship between path length and KB size.   Even if a KB uses a
non-parallelizable ontology language this doesn't matter much if the paths
that cause problems are short compared to the size of the KB.  The
non-parallel aspects of path processing of short paths will not be noticable
with a small number of threads (certainly up to hundreds) and maybe not even
for very many threads (because of general overhead).
\end{quote}

We added xxx to adress this...

\begin{quote}
  The second sentence of Section 3 is badly stated and misleading.  It is definitely true that for a Datalog program $\langle R, I \rangle$ the rule set $R$ is fixed.  However, the initial facts $I$ are also fixed.
\end{quote}

We did ...

\begin{quote}
It is not clear whether the classes of Datalog programs considered in the
paper have to share the same rule set.  The discussion previous to
Definition 1 seems to so indicate, but there is nothing in the definition
that says so.
\end{quote}

We did...

\begin{quote}
There is no definition of what the parents of a node in a graph are.
However, it is quite obvious what the definition should be.
\end{quote}

...

\begin{quote}
Example 6 does need a bit of revision.  It is not the case that Oex3 cannot
be materialzed tractably in parallel.  Instead there has to be a class of
ontologies that cannot be so materialized.  
\end{quote}

...

\begin{quote}
It would be much better if the ontology names corresponded to the example
numbers, i.e., the ontology in Example 6 was Oex6.
\end{quote}

...

\begin{quote}
The notion of a simple concept is very much tied to the algorithm developed
in the paper.  How often do actual ontologies have constructs that could
make them intractable that are rended tractable by having simple concepts?
If there are none, then this class is uninteresting, particularly as the
paper makes such strong claims about worrying about particular ontologies.
This is different from an ontology only containing simple concepts.
\end{quote}

...

\begin{quote}
The paper is all about parallel tractability of materialization, both
theoretical and practical.  However the paper gives insufficient attention
to RDFS, which is the main ontology language used in practice, and
particularly for large knowledge bases.  Even though RDFS is laughably
simple it does not have parallel tractability of even entailment.  The paper
should better examine parallel tractability in the context of RDFS or RDFS
with minor extensions.

Of course, RDFS does not fit into the approach of the paper, as it does not
a-priori separate classes and individuals.  The paper needs to discuss this
issue, as it prominently mentions RDFS.   
\end{quote}

...

\begin{quote}
The paper does a bad job of describing why materialization in RDFS is not
parallizable.  RDFS KBs have an infinite number of consequences, so it is
certainly not possible to materialize all of them quickly.  There are,
however, well known tricks that can be used to generate a finite
representation of this infinite set of consequences.  However, even this
cannot be done quickly in parallel.  In fact, even RDFS entailment cannot be
done quickly in parallel, because RDFS allows for a kind of twisted
interaction (which I think it similar to twisted paths in the paper) when
computing subproperty relationships.  Of course, having subproperties of
rdfs:subPropertyOf is not a usual thing to do in RDFS.  The paper needs to
be much more clear on its relationship to RDFS.
\end{quote}

...

\begin{quote}
The paper mentions YAGO early on but does not indicate why YAGO admits good
performance for parallel reasoning.  YAGO uses a very simple ontology
language - RDFS plus two simple extensions.  It is thus a prime candidate
for close examination to show how practical results differ from worst-case
theoretical ones.  However, the paper does not have any examination of
reasoning on YAGO.  Such an examination is very much called for.  Similarly
an examination of the performance of other large RDFS-based KBs, such as
DBpedia, would be very useful.   If YAGO is not benchmarked then it needs to
be dropped entirely.   Similarly for LUBM.
\end{quote}

...

\begin{quote}
The benchmarking part of the paper is very limited and extremely hard to
follow.    Only eight ontologies were benchmarked.  Why were these eight
chosen?  Only considering at most eight threads is much too constricting.  
\end{quote}

...ontologies must satisfy the properties for parallel tractability...

\begin{quote}
More detail is needed on how the KBs were constructed.  Chain length is a
very important aspect of the KBs, but nothing is said of how the instances
in the KB are connected together.  It is not stated which ontologies belong
to which DL.
\end{quote}

...

\begin{quote}
Figure 5 is very misleading because it uses different scales for the two
systems.   There are several immediate questions that arise from this Figure
that are poorly handled in the paper.  First, why does going from one thread
to two make such a difference for ParallelDHL?  Without a convincing
explanation for this surprising aspect of the benchmarking it is hard to
trust any of the rest of the benchmarking.  The other surprising aspect of
Figure 5 is that RDFox speeds up so little.  There is a short explanation of
why this is so, but I would have liked to see a longer one.
\end{quote}

...

\begin{quote}
The computation of the speedup numbers is wholely unexplained.  Why should
this number have any importance?
\end{quote}

...

\begin{quote}
There is a claim that the benchmarking distinguishes between the Ddhlo and
D-dhlo.  However, Figure 5 doesn't really show this well, if at all.   For
example, why should processing slow down with more threads at all?  Further,
it appears that even for several of the more well-behaved ontologies that there
is very little speedup between 6 and 8 processors.
\end{quote}

...

\begin{quote}
The benchmarking should be redone with more ontologies and with more
threads.  As well, there needs to be better analysis of where parallelism is
failing.  For example, if chains are short, then there should be no problem
in achieving near-perfect thread utilization, ignoring memory contention
problems.
\end{quote}

\begin{quote}
Wording and structural problems:

The paper has so many grammatical errors near its beginning that I gave up
marking them after the first two pages.  I give a few that I noticed here.\\
- [high-performance] computing platforms [] to\\
- do exist ontologies[] for which materialization\\
- (we use experiments to show [this] in Section 6).\\ 
- There has been work [] propos[ing] parallel reasoning algorithms and employ[ing] parallel\\
- e.g., dictionary encoding, balanc[ing] workload, [and] data partitioning,\\
- It has also been shown that[] some well-known large-scale ontologies,\\
The paper needs to be carefully copy-edited before any publication to remove
these errors.

"parallely tractable" is very grating.  It would be better to rewrite
sentences to not use it.
\end{quote}

TODO: Remove the quote above once the issues are fixed.

The paper has been carefully proof-read and hopefully all wording and structural issues are fixed now.

\begin{quote} 
The abstract of the paper is much too long.  An abstract is supposed to be
one not-long paragraph.  Instead it is here two long paragraphs.  The
abstract needs to be cut down to a suitable size before any publication.
\end{quote}

TODO: shorten the abstract.

Indeed, the abstract has been shortened significantly now.

\section{Rebuttal to Review 2}

\begin{quote}
First, in the discussion under Lemma 1, it says $|P^*|$ is polynomial in the size of $P$, which is imprecise, as $|P^*|$ is polynomial in the size of $I$ but not necessarily in the size of $R$. This is not a big issue if one considers NC for data complexity. What is less straightforward and more of concern is the statement that Step~2 costs only one time unit. As Step~2 involves checking the applicability of $B_1, \ldots, B_n \to H$ in $G$ and updating $G$, it depends on the size of $G$ and the size of $G$ is not a constant w.r.t.\ that of $I$. Footnote~4 suggests checking applicability of the rule can be done in constant time via an index. What about updating $G$ and maintaining the index?
\end{quote}

...

\begin{quote}
The computation of the rch relation $S_{\text{rch}}$ ($\dag$) looks problematic, as it requires ``H has been added to G''. If it were the case, then in Example~4 the $S_{\text{rch}}$ would always be empty, as none of the rule heads $H$ would be added to $G$. Again, the complexity analysis of $A_{\text{opt}}$ is unclear and possibly flawed. Step~2 uses an NC algorithm to compute the transitive closure of $S_{\text{rch}}$, which runs in poly-logarithmic time w.r.t.\ the size of $S_{\text{rch}}$ not necessarily so w.r.t.\ the size of $I$. What is the size of $S_{\text{rch}}$ w.r.t.\ that of $I$? Similarly as above, it involves updating $G$, which depends on the sizes of both $G$ and $S_{\text{rch}}$. Would it be still in poly-logarithmic time? 
\end{quote}

...

\begin{quote}
Finally, the algorithms assume each processor handling a rule instantiation. I wonder how it is implemented in ParallelDHL.
\end{quote}

...

\begin{quote}
I list some relatively minor issues regarding presentation and technical details:

- p5, saying ``y is a free variable'' in $A(x) \to \exists y.R(x,y)$ is confusing\\
-  p6, the definition of (complete) materialisation graphs seems problematic. According to Definition~2, a graph containing all ground atoms in $T_R^\omega(I)$ but no edge is a materialisation graph and is complete. In particular,  it does not guarantee ``the nodes whose in-degree is $0$ are the original facts in I'' as claimed.\\
- p6, the use of $\forall$ in Definition~2 is confusing and even worse in p18 Algorithm PRC.\\
- p20, it mentions a group of materialisation rules in YAGO which belong to $D_{\text{dhl}}$. It deserves a clarification on what these rules are like and why they do not fall into existing parallel tractable classes.\\
- p20, Table 2, it should mention how these ontologies were selected. \\
- p21, ``devided'' —> ``divided''\\
- p21, the discussion on line graph lg1 is confusing (i.e.\ ``line-2 stays higher than line-6, but lower than line-4''). Maybe I was looking at the wrong place. \\
- p25, the section heading ``Reference'' occur twice.
\end{quote}

TODO: remove the quote once the issues are fixed.

All minor issues have been fixed.

\end{document}